---
title: "Students favorite places in Denmark"
author: "Astrid Hedegaard Lund, Katrine Foldager, Lasse Axel Lundgreen"
date: "2026-05-13"

output:
  html_document:
    toc: true
    toc_float: true
    theme: cosmo
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(leaflet)
library(tidyverse)
library(googlesheets4)
library(htmlwidgets)
library(htmltools)
```

# Introduction
In this portfolio project, we are analyzing a dataset containing a list of the students of DAM spring 2026 favorite places in Denmark.

We will be creating interactive maps using Leaflet, to answer the following research question:
**Are there any trends in the students of DAM spring 2026 favorite places in Denmark?**

Furthermore this project demonstrates:
- How to create interactive maps with Leaflet, as well as adding esri layers. 
- How to import data from Google Sheets.
- Create marker clustering and popups with descriptions.
- Saving Leaflet maps as HTML files.

## Task 1: Creating a basic Leaflet map og Denmark

```{r danmap, include=TRUE}
l_dk <- leaflet() %>%
  setView(10.85089, 55.2339084, zoom = 6) %>%
  addTiles()

esri <- grep("^Esri", names(providers), value = TRUE)

for (provider in esri) {
  l_dk <- l_dk %>%
   addProviderTiles(
      provider = provider,
      group = provider
    )
}
  
DANmap <- l_dk %>%
  addLayersControl(
    baseGroups = esri,
    options = layersControlOptions(collapsed = FALSE)
  )

DANmap
```

## Task 2: Importing data from Google Sheets

```{r data, include=TRUE}
# Uncomment if authentication issues occur
# gs4_deauth()

places <- read_sheet(
  "https://docs.google.com/spreadsheets/d/1PlxsPElZML8LZKyXbqdAYeQCDIvDps2McZx1cTVWSzI/edit#gid=124710918",
  col_types = "cccnncnc",
  range = "DAM2026"
)

glimpse(places)
```
### Question
**Are the Latitude and Longitude columns present and numeric?**
Yes, both columns are present and contain decimal degree coordinates.

## Task 3: Creating a map with individual plots

```{r Favorite_places, include=TRUE}
Favorite_places <- leaflet() %>%
  addTiles()
  
esri <- grep("^Esri", names(providers), value = TRUE)

for (provider in esri) {
  Favorite_places <- Favorite_places %>%
   addProviderTiles(
      provider = provider,
      group = provider
    )
}
  
Favorite_places <- Favorite_places %>%
  addMarkers(
    lng = places$Longitude,
    lat = places$Latitude,
    popup = paste(
     "<strong>Place:</strong>",
      places$Placename,
      "<br><strong>Description:</strong>",
      places$Description,
      "<br><strong>Type:</strong>",
      places$Type
    ),
    group = "Favorite Places"
  ) %>%
  addLayersControl(
    baseGroups = esri,
    overlayGroups = "Favorite Places",
    options = layersControlOptions(collapsed = FALSE)
  )

Favorite_places
```

## Task 4: Creating a map with clustered plots

```{r clustering, include=TRUE}
Favorite_places_cluster <- leaflet() %>%
  addTiles()

esri <- grep("^Esri", names(providers), value = TRUE)

for (provider in esri) {
  Favorite_places_cluster <- Favorite_places_cluster %>%
   addProviderTiles(
      provider = provider,
      group = provider
    )
}
  
Favorite_places_cluster <- Favorite_places_cluster %>%
  addMarkers(
    lng = places$Longitude,
    lat = places$Latitude,
    clusterOptions = markerClusterOptions(),
    popup = paste(
      "<strong>Place:</strong>",
      places$Placename,
      "<br><strong>Description:</strong>",
      places$Description,
      "<br><strong>Type:</strong>",
      places$Type
    ),
    group = "Favorite Places"
  ) %>%
  addLayersControl(
    baseGroups = esri,
    overlayGroups = "Favorite Places",
    options = layersControlOptions(collapsed = FALSE)
  )

Favorite_places_cluster
```

## Task 5: Saving the maps as HTML

```{r save, include=TRUE}
saveWidget(Favorite_places, "Favorite_places.html", selfcontained = TRUE)

saveWidget(Favorite_places_cluster, "Favorite_places_cluster.html", selfcontained = TRUE)
```

# Reflection: Pro's and cons for each map
**Without Clustering**
- **Pros:** Easier to immediately identify exact locations and overall trends.
- **Cons:** Can feel visually cluttered and overwhelming.

**With Clustering**
- **Pros:** Visually cleaner and easier to navigate.
- **Cons:** Requires zooming for detailed inspection

# Conclusion
See our portfolio document for our conclusion on our research question.